---
title: "Start"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: FALSE
    code_folding: show
    number_sections: TRUE
---

# SET-UP

```{r markdown, echo=F}
library(knitr)
opts_chunk$set(echo = T, collapse = T, fig.align = "center", fig.width = 9, fig.height = 6)
options(width = 300)
```    

```{r settings, message = F}
source('../../0.0_settings.R')

this.year <- 2024
```

# Input data
## Read
```{r read, message = F}
new <- T
if(new){
    bio <- get.bio(species='maquereau',user=imlp.user,password=imlp.pass)
    save(bio,file=paste0('../../Rdata/bio_',Sys.Date(),'.Rdata'))
}else{
    df <- file.info(list.files("../../Rdata/", full.names = T,pattern="bio_"))
    load(rownames(df)[which.max(df$mtime)])
}

```

## Subsetting and outlier removal
Could do outlier removal better.
```{r outliers, message = F}
# subset
bio.waa <- bio[!is.na(bio$weight) & 
               !is.na(bio$agef) & 
                bio$agef>0 & 
                bio$agef<21,
               c('year','month','date','gear','nafo','weight','agef','sample.id','length.frozen')]
names(bio.waa)[7] <- 'age'
names(bio.waa)[9] <- 'length'

# remove outliers (see plot)
bio.waa$outlier <- FALSE
bio.waa$outlier <- ifelse(bio.waa$weight<0.03,TRUE,bio.waa$outlier)              # less than 30gr
bio.waa$outlier <- ifelse(bio.waa$age>17,TRUE,bio.waa$outlier)                   # at least 4 out of 6 age 18 have wrong length AND weight
bio.waa[!bio.waa$outlier,] <- ddply(bio.waa[!bio.waa$outlier,],c('age'),transform,
                                    outlier=ifelse((weight<=quantile(weight,0.0001)|weight>=quantile(weight,0.9999)),TRUE,outlier))  # quick and dirty way to remove extremes

# cleaning
bio.waa$age.group <- ifelse(bio.waa$age < 10, bio.waa$age, 10)
bio.waa$length <- roundFish(bio.waa$length*100,0.5) # round to half centimeters
```

## plots {.tabset}

### age vs weight
```{r plot age_weight}
ggplot(bio.waa,aes(x=age,y=weight,col=outlier))+
    geom_point()
```

# Calculations
## Try different options 
Even trial run with different selectivity gears (despite not a great idea).
```{r options}
bio.waa <- bio.waa[!bio.waa$outlier,]
bio.waa$outlier <- NULL

# 4T + gillnets + june/july
df1 <- bio.waa[bio.waa$month %in% c(6,7) &
               bio.waa$nafo %in% c('4T') &
               bio.waa$gear %in% c("GND", "GNS", "GN"),]

waa1 <- get.waa(df1$year,df1$age.group,df1$weight,df1$sample.id)
waa1$option <- '4T_gill_JunJul'

# 4TVWX + gillnets + june/july
df2 <- bio.waa[bio.waa$month %in% c(6,7) &
               bio.waa$nafo %in% c('4T', "4V", "4W", "4X") &
               bio.waa$gear %in% c("GND", "GNS", "GN"),]

waa2 <- get.waa(df2$year,df2$age.group,df2$weight,df2$sample.id)
waa2$option <- '4TVWX_gill_JunJul'

# 4T + gillnets/line+jig + june/july
df3 <- bio.waa[bio.waa$month %in% c(6,7) &
               bio.waa$nafo %in% c('4T') &
               bio.waa$gear %in% c("GND", "GNS", "GN","LHM","LMP","LX","LHP"),]

waa3 <- get.waa(df3$year,df3$age.group,df3$weight,df3$sample.id)
waa3$option <- '4T_gillhandjig_JunJul'

# 4Tvwx + gillnets/line+jig + june/july
df4 <- bio.waa[bio.waa$month %in% c(6,7) &
               bio.waa$nafo %in% c('4T', "4V", "4W", "4X") &
               bio.waa$gear %in% c("GND", "GNS", "GN","LHM","LMP","LX","LHP"),]

waa4 <- get.waa(df4$year,df4$age.group,df4$weight,df4$sample.id)
waa4$option <- '4TVWX_gillhandjig_JunJul'
```

## Select option 
Going with  4TVWX + gillnets + june/july
```{r select}
thisbio <- df2
thiswaa <- waa2   
```

## plots {.tabset}

### waa
```{r plot comparison_waa}
waa.comp <- rbind(waa1,waa2,waa3,waa4)
ggplot(waa.comp, aes(x=year, y=w, col=as.factor(age))) + 
    geom_line(linewidth=1) + 
    facet_wrap(~option)+
    labs(x='Year',y='Weigt (kg)',col='Age')+
    theme_minimal(base_size = 14) 
```

### number of fish
```{r plot comparison_nfish}
ggplot(waa.comp, aes(x=year, y=age)) + 
    geom_tile(aes(fill=n)) + 
    geom_text(data=waa.comp[waa.comp$n<10,],aes(label=n))+
    facet_wrap(~option)+
    labs(x='Year',y='Age',fill='n')+
    theme_minimal(base_size = 14)+
    scale_fill_viridis_c(direction = -1)+
    scale_x_continuous(expand=c(0,0))
```

### number of samples
```{r plot comparison_Nsample}
ggplot(waa.comp, aes(x=year, y=age)) + 
    geom_tile(aes(fill=N)) + 
    geom_text(data=waa.comp[waa.comp$N<5,],aes(label=N))+
    facet_wrap(~option)+
    labs(x='Year',y='Age',fill='N')+
    theme_minimal(base_size = 14)+
    scale_fill_viridis_c(direction = -1)+
    scale_x_continuous(expand=c(0,0))
```

# How many fish do we minimally need?
## Simulation
```{r plot fish_min}
newm <- T
if(!newm){
    df <- file.info(list.files("../../Rdata/", full.names = T,pattern="sims_"))
    load(rownames(df)[which.max(df$mtime)])
}else{
    waa.tot <- ddply(thisbio,c('age.group'),summarise,w=mean(weight),wsd=sd(weight),n=length(weight))
    re <- function(test,true) (test-true)/true
    
    sims <- ldply(1:1000,function(i){
        ldply(unique(thisbio$age.group),function(a){
            ldply(1:100,function(f){
                  # sample from all ages (across all years to have enough to sample from; might not be the case for age 1)
                  d <- thisbio[thisbio$age.group==a & !is.na(thisbio$age.group),]
                  
                  # Sample 1 to 100 fish weights for each step
                  ws <- sample(d$weight, f, replace = TRUE)
                  
                  # Calculate mean weight of fish sampled
                  wav <- mean(ws)
                  
                  # Estimate relative error
                  RE <- re(wav,waa.tot[waa.tot$age.group==a,'w'])
                  
                  return(c(sim=i,age=a,n=f,RE=RE))
            })
        })
    })

save(sims,file=paste0('../../Rdata/sims_',Sys.Date(),'.Rdata')) 
}

check <- ddply(sims,c('age','n'),summarise,mean=mean(RE),low=mean(RE)-1.96*sd(RE),high=mean(RE)+1.96*sd(RE))

pd <- position_dodge(0.1) 
plots <- lapply(unique(check$age),function(x){
    ggplot(check[check$age==x,], aes(x = n, y = mean)) +
        geom_line(size = 1.5) +
        geom_errorbar(size = 0.5, aes(ymin=low, ymax=high), width=1.5, position=pd) +
        geom_line(position=pd, size=1) +
        geom_point(position=pd, size=1) +
        geom_hline(yintercept=0.1, linetype="dashed", color = "red") +
        geom_hline(yintercept=-0.1, linetype="dashed", color = "red") +
        theme_minimal(base_size = 14) + 
        ylab("Relative error with 95%CI") + 
        xlab("Number of fish sampled") +
        scale_x_continuous(breaks = seq(0,100,5)) +
        ggtitle(paste("Age", x, sep = " "))
})


#Remove year-age group that contain less than 10 fish 
```

## plots {.tabset}
Note that this is an easy way to determine error in function of the number of fish measured. It might be overestimated because the true distribution of values (weights) is lower than assumed here (by grouping all years).

### age1
```{r plot re_age1}
plots[[1]]
```

### age2
```{r plot re_age2}
plots[[2]]
```

### age3
```{r plot re_age3}
plots[[3]]
```

### age4
```{r plot re_age4}
plots[[4]]
```

### age5
```{r plot re_age5}
plots[[5]]
```

### age6
```{r plot re_age6}
plots[[6]]
```

### age7
```{r plot re_age7}
plots[[7]]
```

### age8
```{r plot re_age8}
plots[[8]]
```

### age9
```{r plot re_age9}
plots[[9]]
```

### age10
```{r plot re_age10}
plots[[10]]
```


# What is the effect of working with subsample data?
Scale by catch length-frequencies. 

## Read
```{r read_lf, message = F}
newf <- FALSE
if(newf){
    lf <- get.lf(species='maquereau',user="pec_iml_ro",password="Pecimlro_1!")
    save(lf,file=paste0('../../Rdata/lf_',Sys.Date(),'.Rdata'))
}else{
    df <- file.info(list.files("../../Rdata/", full.names = T,pattern="lf_"))
    load(rownames(df)[which.max(df$mtime)])
}

```

## Proportions
```{r prop_lf, message = F}
props <- lf[!is.na(lf$length) &
            lf$month %in% unique(thisbio$month) &
            lf$nafo %in% unique(thisbio$nafo) &
            lf$gear %in% unique(thisbio$gear),]

props <- ddply(props,c('year','length'),summarise,n=length(length))
props <- ddply(props,c('year'),transform,prop=n/sum(n))

props$length <- roundFish(props$length*100,0.5)

propsall <- expand.grid(year=unique(thisbio$year),length=seq(min(thisbio$length,na.rm=TRUE),max(thisbio$length,na.rm=TRUE),by=0.5))
propsall <- merge(propsall,props,all.x=TRUE)
propsall <- ddply(propsall,c('year'),transform,prop=zoo::na.locf(zoo::na.locf(prop,na.rm = FALSE), fromLast = TRUE)) # not great, but little impact

thisbiolf <- merge(thisbio,propsall,all.x=TRUE)
thisbiolf[is.na(thisbiolf$prop),'prop'] <- max(thisbiolf$prop,na.rm = T) # not great, but nearly all for same year so no impact

waa5 <- ddply(thisbiolf,c('year','age.group'),summarise,
          w=weighted.mean(weight,prop,na.rm=T),
          wsd=NA, # see Hmish package
          n=length(weight),
          N=length(unique(sample.id)))
waa5$option <- 'with lf'
names(waa5)[2] <- 'age'
```

## plots {.tabset}

### comparison pairwase
```{r plot waalf}
waa.comp2 <- rbind(thiswaa,waa5)
ggplot(waa.comp2, aes(x=year, y=w, col=as.factor(age))) + 
    geom_line(size=0.5) + 
    facet_wrap(~option)+
    labs(x='Year',y='Weigt (kg)',col='Age')+
    theme_minimal(base_size = 14) 
```

### comparison by age
```{r plot waalf2}
ggplot(waa.comp2, aes(x=year, y=w, col=option)) + 
    geom_line(size=0.5) + 
    facet_wrap(~age,scale='free_y')+
    labs(x='Year',y='Weigt (kg)',col='Age')+
    theme_minimal(base_size = 14) 
```

### length props (facets)
```{r plot lengthprops}
ggplot(propsall,aes(x=length,y=prop))+
    geom_line(size=0.5)+
    facet_wrap(year~.)
```

### length props (superposed)
```{r plot lengthprops2}
ggplot(propsall,aes(x=length,y=prop,col=year,group=year))+
    geom_line(size=0.5)
```

# What if I would have used a model?

Not included. Could predict waa (and impute missing values) based on model predictions. However, need interactions (year*gear), nonlinear effects (year), etc. On a dataframe with millions of lines it is extremely slow. Appears overly complex and not very practical.

Test code removed.

# Fill and smooth 

Fit Noel Cadigan's state-space model for waa.
1) filter out noise.
2) fill in gaps.

Fits an AR1 process in 3 directions (age, year, cohort). Available in catchR package.

## fit model
```{r waa_fit, message = F, fig.height = 11,fig.width = 10,results = 'hide',message=FALSE}
waam <- thiswaa[thiswaa$n>=10,]
waam$cv <- with(waam,wsd/w)

all <- expand.grid(year=1968:this.year, age=min(waam$age):max(waam$age))
waam <- merge(all, waam, all.x = TRUE)

waa.fit <- armatrix.fit(year=waam$year,age=waam$age,x=waam$w,cv=waam$cv,shrink.cv = 0)
waa.fit  # convergence ok

save(waa.fit,file=paste0("../../Rdata/fit.armatrix/",this.year,"base_cv1shrink0.Rdata"))
```

## plots {.tabset}

### waa
```{r waa_final, message = F, fig.height = 3,fig.width = 5,results = 'hide',message=FALSE}
p<- armatrix.waa(waa.fit)+scale_color_viridis_d()+scale_y_continuous(limits=c(0,0.9),expand=c(0,0))

p + labs(x="Année | Year", y= "Poids | Weight (kg)", col="Âge | Age")

```

### Predicted obs
```{r waa_predobs, message = F, fig.height = 6,fig.width = 11,results = 'hide',message=FALSE}
p<- armatrix.predobs(waa.fit,scale='free',ncol=4)

p+ ggtitle("Prédit vs Observé | Predicted vs Observed") +labs(x="Année | Year", y="Poids à l'âge | Weight-at-age (kg)")

```

### Predicted
```{r waa_pred, message = F, fig.height = 10,fig.width = 10,results = 'hide',message=FALSE}
armatrix.pred(waa.fit,scale='free',ncol=3)
```

### Resid bubble
```{r waa_res, message = F, fig.height = 3,fig.width = 10,results = 'hide',message=FALSE}
armatrix.res(waa.fit)
```

### Resid dot
```{r waa_res2, message = F, fig.height = 5,fig.width = 10,results = 'hide',message=FALSE}
   d <- waa.fit$output
    d$cohort <- with(d, year - age)
    p1 <- ggplot(d, aes(x = year, y = res)) + geom_text(aes(label = age), 
        size = 2) + geom_hline(yintercept = 0, linetype = "dashed") + 
        labs(y = "", x = "Année | Year") + geom_smooth()
    p2 <- ggplot(d, aes(x = cohort, y = res)) + geom_point(size = 1) + 
        geom_hline(yintercept = 0, linetype = "dashed") + labs(y = "", 
        x = "Cohorte | Cohort") + geom_smooth()
    p3 <- ggplot(d, aes(x = age, y = res)) + geom_point(size = 1) + 
        geom_hline(yintercept = 0, linetype = "dashed") + labs(y = "", 
        x = "Âge | Age")
    p4 <- ggplot(d, aes(x = pred_exp, y = res)) + geom_point(size = 1) + 
        geom_hline(yintercept = 0, linetype = "dashed") + labs(y = "", 
        x = "Prédites | Predicted")
    grid.arrange(p1 , p2, p3, p4, ncol = 2, left = "Résidus standardisés | Standardized residuals")


```

### effects
```{r waa_effects, message = F, fig.height = 10,fig.width = 10,results = 'hide',message=FALSE}
armatrix.effects(waa.fit) # empty plot because of ggplotGrob function within (used to align plots)
```

### cvs
```{r waa_cv, message = F, fig.height = 10,fig.width = 10,results = 'hide',message=FALSE}
armatrix.cvs(waa.fit,scale='free',ncol=3)
```
# Save results
```{r waa_result, message = F}
waaf <- waa.fit$output
waaf <- waaf[,c(1:2,9)]
names(waaf)[3] <- 'weight'

s <- dcast(waaf,year~age,value.var = 'weight')
s[,2:ncol(s)] <- round(s[,2:ncol(s)] ,3)
write.csv(s, file=paste0('../../csv/waa_',this.year,'_base_cv1shrink0.csv'),row.names = FALSE)

p <- armatrix.waa(waa.fit,ylab='Weight (kg)')+scale_color_viridis_d()
ggsave(filename = paste0('../../img/waa',this.year,'.png'),plot = p,units = 'cm',height = 8,width = 14)
```